---
title: "StatisticalInferenceProject"
author: "Alex Mihailov"
date: "2024-09-04"
output:
  pdf_document: default
  html_document: default
---

## Steps

1. Show the sample mean and compare it to the theoretical mean of the distribution.
2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.

## Data

```{r}
library("data.table")
library("ggplot2")

set.seed(100)

lambda <- 0.2

n <- 40

simulations <- 1000

simulated_exponentials <- replicate(simulations, rexp(n, lambda))

means_exponentials <- apply(simulated_exponentials, 2, mean)
```

## Question 1

Show where the distribution is centered at and compare it to the theoretical center of the distribution

```{r}
analytical_mean <- mean(means_exponentials)
analytical_mean
```
Mean
```{r}
theory_mean <- 1/lambda
theory_mean
```
visualization
```{r}
hist(means_exponentials, xlab = "mean", main = "Exponential function simulations")
abline(v = analytical_mean, col = "red")
abline(v = theory_mean, col = "blue")
```

The analytics mean is 4.999702 the theoretical mean 5. The center of distribution of averages of 40 exponentials is very close to the theoretical center of the distribution.


## Question 2

Show how variable it is and compare it to the theoretical variance of the distribution

Standard deviation of distribution
```{r}
standard_deviation_dist <- sd(means_exponentials)
standard_deviation_dist
```

Standard deviation from analytical expression
```{r}
standard_deviation_theory <- (1/lambda)/sqrt(n)
standard_deviation_theory
```
Variance of distribution
```{r}
variance_dist <- standard_deviation_dist^2
variance_dist
```
Variance from analytical expression
```{r}
variance_theory <- ((1/lambda)*(1/sqrt(n)))^2
variance_theory
```

Standard Deviation of the distribution is 0.8020251 with the theoretical SD calculated as 0.7905694. The Theoretical variance is 0.6432442. The actual variance of the distribution is 0.625

## Question 3
Show that the distribution is approximately normal

```{r}
xfit <- seq(min(means_exponentials), max(means_exponentials), length=200)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(n)))
hist(means_exponentials,breaks=n,prob=T,col="yellow",xlab = "means",main="Density of means",ylab="density")
lines(xfit, yfit, pch=30, col="brown", lty=5)
```

```{r}
qqnorm(means_exponentials)
qqline(means_exponentials, col = 2)
```

Due to Due to the central limit theorem (CLT), the distribution of averages of 40 exponentials is very close to a normal distribution.