---
title: "StatisticalInferenceProject2"
author: "Alex Mihailov"
date: "2024-09-04"
output:
  pdf_document: default
  html_document: default
---

Overview
The project aim is to analyze the ToothGrowth data in the R datasets package.


1. Load the ToothGrowth data and perform some basic exploratory data analyses
```{r}
library(ggplot2)
library(tinytex)
library(datasets)

data(ToothGrowth)
str(ToothGrowth)
head(ToothGrowth, 5)
tail(ToothGrowth, 5)

summary(ToothGrowth)
```


2.Basic summary of the data
```{r}
Supplement_mean = split(ToothGrowth$len, ToothGrowth$supp)
sapply(Supplement_mean, mean)
```

Graph
```{r}
ggplot(aes(x=supp, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=supp)) + 
    xlab("Supplement Type") +
    ylab("Tooth length") 
```

3. Using confidence intervals to compare growth of tooth by supplement dose
```{r}
unique(ToothGrowth$dose)
```

There are 3 groups: 0.5, 1, and 2
Graph shows relationship between Tooth length to Dose
```{r}
ggplot(aes(x = factor(dose), y = len), data = ToothGrowth) + 
    geom_boxplot(aes(fill = factor(dose))) +
    labs(title="Tooth Lenght relationship to Dosage")

```
T-test for dose 0.5 mg:
```{r}
t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == .5, ])
```
T-test for dose 1 mg:
```{r}
t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 1, ])
```
T-test for dose 2 mg:
```{r}
t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 2, ])
```
Conclusion:

For all three dosages, the p-value of this test is is less than 0.5, a evidence that we can reject the null hypothesis. We can infer that supplement type has no effect on tooth growth, and increasing the dose level leads to increased tooth growth.

